Syntactically Informed Natural Language Inference

### 2019-03-19 cse517/report/main.tex
Added model changes

\begin{document}
\maketitle
\begin{abstract}
  Abstract.
\end{abstract}

@@

Natural language inference (NLI) is the task of characterizing entailment and
contradiction relationships between texts. We seek to improve models which
perform NLI by adding syntactic features into existing models. In particular,
we will primarily investigate the well-known decomposable attention (DA) neural
network \citep{Parikh2016-em} which obtained state of the art results on the
SNLI \citep{Bowman2015-is} dataset at its time of publication while using
drastically fewer parameters than previous NLI models.

@@

We extend the DA model by encorporating syntactic features.
\dpcomment{Include brief summary of constributions here.}
We primarily
evaluate on the domain specific Scitail \citep{Khot2018-th} for both its
smaller size and lack of annotation artifacts when compared to SNLI
\citep{Gururangan2018-lj}.

@@

Our project will build on previous work done by \citet{Pang2018-syntail}.

@@

>>>>

\begin{document}
\maketitle
\begin{abstract}
This project explores methods for improving performance on the task of natural language inference by incorporating syntactic information. Specifically we modify the decomposable attention neural network \citep{Parikh2016-em} by incorporating encoded syntax parses of the input statements at different layers of the network. We supplement these experiments by using different syntax parsers including pre-trained constituency and dependency parsers as well as a constituency parser trained separately for this task. 
\end{abstract}

@@

Natural language inference (NLI) is the task of characterizing entailment and
contradiction relationships between texts. We improve NLI models by adding syntactic
features into existing models.
We will primarily investigate the well-known decomposable attention (DA) neural
network \citep{Parikh2016-em}

@@

We extend the decomposable attention (DA) model by \citet{Parikh2016-em}
incorporating syntactic features.
\dpcomment{Include brief summary of contributions here.}
The DA model obtained state-of-the-art results on the
SNLI \citep{Bowman2015-is} dataset at its time of publication while using
drastically fewer parameters than previous NLI models.
We primarily
evaluate on the domain specific Scitail \citep{Khot2018-th} for both its
smaller size and lack of annotation artifacts when compared to SNLI
\citep{Gururangan2018-lj}.

@@

@@

\section{Models}

The syntactic entailment model is derived from the decomposable attention model for natural language inference described by \citet{Parikh2016-em}. In the initial iteration of the network, syntax is not incorporated until the very last feed-forward layer. This means that syntax is not considered when calculating attention between the premise and the hypothesis. We speculated that by introducing syntax to the network prior to calculating attention we would see better results as the syntax would help influence the attention calculations and reveal the most important phrases of each input statement.

In the ``V3" model, we perform the same initial step of embedding the premise and hypothesis and expanding their dimensions (through a fully-connected layer) prior to attention. However, instead of calculating our similarity matrix using solely the embedded hypothesis and premise vectors as input, we also provide the encoded syntax parses of both inputs. The concatenation of these four vectors is used to derive the similarity matrix which is then utilized in the attention computations between the premise and hypothesis.

With the ``V5" model we tried to incorporate the computed syntax of the inputs even earlier. Instead of concatenating the encoded syntax parses with the embedded inputs and using that as input for the computation of the similarity matrix, we concatenated the syntax parses with their respective inputs and passed that through the first fully-connected layer. The two resulting vectors representing the embedded inputs with syntactic information are then used to compute the similarity matrix.

Our ``V7" model took an additional step of learning weights for the encoded syntax parses by passing the vectors through an additional fully-connected layer before introducing them into the rest of the network. This model was built on top of the ``V3" model such that the weighted syntax parses are concatenated with the embedded premise and hypothesis statements when deriving the similarity matrix.

Supplemental variations were done on top of these restructures of the network. One of the variations that we introduced was a retrained constituency parser. By retraining the constituency parser on the Penn Treebank dataset \citep{marcus1993building} we could change the dimensions of the encoded syntax vectors to better fit with our models. Additionally, we experimented with training our models using the dependency parser model derived by \citet{dozat2016deep} to see if the different types of encoded parses influenced the performance of the network.
